Distributed Topology Discovery in Self-Assembled Nano Network on Chip

### 2014-02-18 PAPER/introduction.tex
Fixed typos, hilighted some doubts

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources deadlock makes them not suitable for
the specific DNA self assembled networks we are assuming as
background.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channels but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
The hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present an approach the solves these
limitations by allowing turn restrictions to be placed locally,
independently from other restrictions. The whole network is
partitioned into segments, and a bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom while preserving network connectivity. This \emph{locality
independence} property, avoiding the requirement of any particular tree/root
choice, would make it the best choice for the given scenario;
however, it still requires the knowledge of the
whole network graph in order to find the segments.

The \disr{} approach presented in this work is a first attempt of
achieving the same goals of a segment based turn prohibition  without
requiring any topology graph or centralized execution. In particular
\disr{} implements segment discovery and turn prohibition using a
completely different mechanism, based on a distributed exchange of
small setup packets requiring on average $15\%$ of the available
resource budget. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of belonging to some segment,
ignoring the presence of other nodes in the same segment and even the
presence of other segments in the network
\end{itemize}

>>>>

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources available to prevent deadlocks makes them not suitable for
the DNA self-assembled networks we are assuming as
scenario.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channel but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploit the creation of a spanning tree of the
topology, placing bidirectional restrictions \hl{by avoiding a packet
to traverse the same link in both up and down directions}.
The hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present an approach that solves these
limitations setting turn restrictions locally,
independently from other restrictions. The whole network is
partitioned into segments, and a bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom while preserving network connectivity. This \emph{locality
independence} property, removing the requirement of choosing a particular tree/root,
would make this approach the best choice for the given scenario;
however, it still requires the knowledge of the
whole network graph in order to find the segments.

The \disr{} approach presented in this work is a first attempt to
achieve the same goals of a segment based turn prohibition without
requiring neither the knowledge of a topology graph nor centralized execution. In particular
\disr{} implements segment discovery and turn prohibition using a
completely different mechanism, based on a distributed exchange of
small setup packets requiring on average $15\%$ of the available
resource budget. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of belonging to \hl{a segment | some segments},
ignoring the presence of other nodes in the same segment and also the
presence of other segments in the network
\end{itemize}
